Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-22 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

While the finite-precision extension of Kochen-Specker theorem asserts
the non-existence of certain kind of QIVPMs, we also want to understand
what a QIVPM could be when it exists. In Sec.~\ref{sec:Interval-Uncertainty},
the only non-trivial QIVPM~ $\bar{\mu}_{3}$ is induced by the state~$\proj{\ps}$.
The way we extending a state to a QIVPM will first be generalized
in this section, and introduce a finite-precision extension of the
Born rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. The
Born rule is essential to compute the conventional quantum probability
because any quantum probability measure can be induced by the Born
rule according to Gleason's theorem~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
We then provide a finite precision variant of Gleason's theorem for
the Born QIVPMs, but there are also some general QIVPMs may not be
induced by any quantum probability measure, i.e., no state is consistent
with these kind of QIVPMs.

When we constructed $\bar{\mu}_{3}$, we replaced the real-valued
probability~$0$ and $\frac{1}{2}$ by $\left[0,0\right]$ and $\left[\frac{1}{2},\frac{1}{2}\right]$,
respectively. Replacing numbers by intervals containing them can be
generalized if we carefully pick a set of intervals and a rule to
select the correct interval when intervals are overlapping. For example,
consider the following collection of intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except the intervals
with length~$\left(2n\right)^{-1}$ at each end. Since these intervals~$[\ell,r]$
are overlapping, we define the non-overlapping interior part of a
interval, $\textrm{Int}\left([\ell,r]\right)$, as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}

@@

\yutsung{Not sure what is the best way embedding the condition for
$k$ in the table...}
\end{table}
Hence, given a real-valued probability~$x$, we could easily replace
it with an interval~$\iota_{n}\left(x\right)=[\ell,r]$ if $x\in\textrm{Int}\left([\ell,r]\right)$.

@@

\begin{thm}[Finite-precision Extension of the Born rule~\citep{HOSTunpublished}]For
any state $\rho$ on a three-dimensional Hilbert space, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, where $\circ$ is the function concatenation.\end{thm}\yutsung{Do
we need to summarize the proof?}

\noindent For example, when $n=2$, two states~$\mathbb{1}/3$ and
$\rho_{0}$ can induce two QIVPMs~$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for any projector $P$ in the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed
straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}

>>>>

While the finite-precision extension of Kochen-Specker theorem asserts
the non-existence of certain kind of QIVPMs, we also want to understand
what a QIVPM could be when it exists. A conventional quantum probability
measure~$\mu$ can be easily be constructed using the Born rule when
one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007},
and Gleason's theorem asserts that the only state consistent with
$\mu$ on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Although the Born rule could incorporate the idea of finite precision
easily, there is a QIVPM such that every state can only consistent
with it on some of projectors. However, if we only focus on the QIVPM~$\bar{\mu}$
induced by a state~$\rho$ according to the Born rule, called the
Born QIVPMs, we found the states consistent with $\bar{\mu}$ will
be closer and closer to $\rho$ while shrinking the length of intervals.

The first step toward a finite precision Born rule is to find a way
replacing the computed probability by an interval. This can be done
by carefully picking a set of intervals and a rule to select the correct
interval when intervals are overlapping. For example, consider the
following collection of intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except the intervals
with length~$\left(2n\right)^{-1}$ at each end. To select the correct
interval from these overlapping intervals~$[\ell,r]$, we define
the non-overlapping interior part of a interval, $\textrm{Int}\left([\ell,r]\right)$,
as in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{n}$ and their interior
with an concrete example for $\mathscr{I}_{2}$. Although intervals~$[\ell,r]$
are overlapping, their interior~$\textrm{Int}\left([\ell,r]\right)$
are non-overlapping.}

@@

\yutsung{Not sure what is the best way embedding the condition for
$k$ in the table...}
\end{table}
Then, given a real-valued probability~$x$, we will select a interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

@@

\begin{thm}[Finite-precision Extension of the Born rule]For any
state $\rho$ on a three-dimensional Hilbert space, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, where $\circ$ is the function concatenation.\end{thm}

\noindent Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$,
it is easy to verify $\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)$.
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, either one
of them projects onto the subspace which the other one project onto,
or their complements are orthogonal which reduced the case we discussed~\citep{HOSTunpublished}.
Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is indeed a QIVPM.
For example, when $n=2$, two states~$\mathbb{1}/3$ and $\rho_{0}$
can induce two QIVPMs~$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$, where $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
and $\rho_{0}=\frac{7}{24}\cdot\proj{0}+\frac{1}{3}\cdot\proj{1}+\frac{3}{8}\cdot\proj{2}$.
The interval-valued probability of the induced QIVPMs, $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$
and $\iota_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$,
for any projector $P$ in the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, are computed
straightforwardly in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$ based on $\mu_{\mathbb{1}/3}^{\mathrm{B}}$
and $\mu_{\rho_{0}}^{\mathrm{B}}$ on $\events_{C}$.}
